% Active TMLR appendix. This file is content-only and is included by
% scripts/build_tmlr_paper.sh before the bibliography.

\section{Claim--Evidence Ledger}
\label{app:claim-ledger}

Table~\ref{tab:claim-ledger} fixes the strongest statement licensed by each
completed evidence layer. It is deliberately asymmetric: later negative or
boundary evidence can narrow an earlier claim, but an unfinished workstream
cannot strengthen it.

\begin{table}[h]
\caption{Claim ceiling by evidence layer. ``Pending'' rows are protocol facts,
not empirical results.}
\label{tab:claim-ledger}
\begin{center}
\small
\begin{tabular}{p{0.18\textwidth}p{0.32\textwidth}p{0.39\textwidth}}
\toprule
Evidence layer & Completed observation & Maximum licensed claim \\
\midrule
Qwen development & 60 machine-audited items; typed minus untyped answer score
+0.078, with positive intervals on both machine-disposition strata & A local,
construction-derived typed-control signal; not human validation or model-family
generality \\
RAMDocs development & Two preregistered 350-item rounds fail G-A; Round~2 is
$-0.0029$ against the frozen strongest baseline & No demonstrated end-to-end
advantage on this upstream-labelled external development distribution \\
WS1 frozen attribution & 226 shared errors are uniquely bucketed; all four
registered hypotheses are not supported & A revision-mediated heterogeneous
effect and observed boundary, not an upstream-only or metric-rescue explanation \\
G-P power audit & Estimated WS2 exact-McNemar power is 0.414 for a +0.078 effect
& WS2 is directional; nonsignificance cannot establish absence \\
WS2 family development & Mistral, Gemma, and Llama are 3/3 positive; combined
typed-minus-untyped answer difference is $+0.0645$, G-F passes, and the
independent release audit is valid & Directional reproduction across these
local machine-audited development families; not external or human validation \\
WS3 public development & Protocol and two imports are frozen, but no model
prediction or verified release exists & No external-boundary result may be
inserted yet \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

The machine-readable sources for these rows are
\path{diagnostics/solo_v1},
\path{diagnostics/ramdocs_v2},
\path{diagnostics/attribution_v1/manifest.json}, and
\path{diagnostics/power_v1/manifest.json}, and
\path{diagnostics/family_dev_v1/manifest.json}. The active roadmap fingerprint is
\texttt{91eb3205fe127271bc5f4882025243d9974a711e311ef074fcbde09aa86e7cf7};
it binds the ten frozen facts F1--F10 and is not changed by this appendix.

\section{Registered Attribution Details}
\label{app:attribution}

The WS1 classifier and verifier were frozen at Git commit
\texttt{cded2b7e76067a59294c687fef01ce990a33feef} before the formal inputs were
read. The release verifier binds that commit to the roadmap, classifier,
evidence builder, protocol module, and classification tests. It records zero
model calls, no held-out/test access, and no human-IAA or publication-gold
claim.

\begin{table}[h]
\caption{Earliest-failure assignment for all 226 RAMDocs items missed by both
FAR and Multi-Query RAG. Every item has exactly one bucket.}
\label{tab:failure-buckets}
\begin{center}
\small
\begin{tabular}{lr}
\toprule
Registered bucket & Count \\
\midrule
Retrieval miss & 4 \\
Conflict undetected & 72 \\
Conflict detected, revision wrong & 103 \\
Answer set incomplete & 35 \\
Answer set overfull & 12 \\
Format / exact-match mismatch & 0 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

All four preregistered hypotheses are \texttt{not\_supported}. For
H-upstream, the 179 complete-retrieval items retain a FAR-minus-baseline exact
difference of $-0.011$ with 95\% bootstrap interval $[-0.050,0.028]$. For
H-metric, descriptive collection F1 changes by $-0.003$ with interval
$[-0.020,0.013]$; this alternative metric never reopens G-A. For
H-conflict-shape, conflict detection is lower on RAMDocs than on the Qwen
diagnostic (0.449 versus 0.883), but the detected subset is only $+0.006$ with
an interval crossing zero. For H-component, all 34 positive continuous Qwen
typed-minus-untyped deltas pass through a changed revision, while none occurs
after detection without a changed revision. These facts support the
heterogeneous-mediator interpretation and reject the registered detection-only
prediction.

\section{Power and Outcome-Conditioned Interpretation}
\label{app:power}

The G-P release uses frozen paired discordance rates and records zero model
calls and no held-out/test access. Table~\ref{tab:power} reports the historical
minimum detectable effects (MDEs) and the prospective WS2 designation.

\begin{table}[h]
\caption{Power audit under two-sided exact McNemar testing. MDE is a design
diagnostic, not an acceptable-effect threshold.}
\label{tab:power}
\begin{center}
\small
\begin{tabular}{lrrrr}
\toprule
Design & $n$ & Discordance & MDE@60\% & MDE@80\% \\
\midrule
Qwen typed/untyped & 60 & 0.317 & 0.171 & 0.206 \\
RAMDocs Round~1 & 350 & 0.091 & 0.038 & 0.047 \\
RAMDocs Round~2 & 350 & 0.089 & 0.038 & 0.046 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

For WS2, the registered three-family design contains 180 paired development
observations. At the Qwen discordance rate and a +0.078 target effect, simulated
exact-McNemar power is 0.414, the probability that the family-clustered
bootstrap lower bound exceeds zero is 0.595, and the probability that at least
two of three family directions are positive is 0.929. The protocol therefore
fixes \texttt{directional\_reproduction} as the claim level. There is one
registered run and no Round~2.

The completed release has positive family-level answer differences for
Mistral, Gemma, and Llama ($+0.0528$, $+0.0673$, and $+0.0735$). The combined
difference is $+0.0645$; stratified exact McNemar counts are 31 versus 9
($p=0.000680$), and the family-cluster bootstrap 95\% interval is
$[+0.0528,+0.0735]$. G-F passes, but the low-power designation and all
machine-audited development qualifiers remain binding.

The verified WS2 result closes the C-line family-inconsistency branch. The
eventual paper branch is selected after the independently verified WS3 release:
at least one matched external condition yields the A-line
mechanism-and-boundary paper, whereas null or mixed WS3 evidence yields the
B-line bounded-transfer paper. Neither branch permits end-to-end superiority,
human-gold, or external-blind language under the current evidence.

\section{Reproducibility and Runtime Contract}
\label{app:reproducibility}

The completed Qwen diagnostic runs one inference pass per method over the same
60 development items, with no repeated sampling or best-of-$n$ selection.
Qwen3.5 9B runs through Ollama at temperature zero, thinking disabled, and a
1,200-token output cap. Retrieval uses two documents per query, BM25
($k_1=1.5$, $b=0.75$), BGE-base-en-v1.5 dense scores, 0.4/0.6 sparse/dense
fusion, BGE-reranker-base over 20 candidates, and nli-distilroberta-base. The
recorded host is WSL2 with an NVIDIA RTX 4060 Laptop GPU. Model digest,
configuration hash, implementation hash, predictions, and report hashes are
retained in the public diagnostic bundle.

Every formal run identity covers its YAML configuration, dev-only operational
input, corpus, model digest, and implementation fingerprints. Checkpoints are
written after each sample, and resume rejects identity drift. Structured model
outputs are schema-validated; provider failures remain errors rather than
silently changing the experimental arm.

The active paper can be rebuilt with
\texttt{bash scripts/build\_tmlr\_paper.sh}. The builder pins the official TMLR
style at commit
\texttt{7bf90efe3a0debbba703c05c43f3ff7e4d4a2992}, records the main-text and
appendix SHA-256 values, and generates only ignored build artifacts. The WS1,
WS2, WS3, G-P, and repository evidence verifiers remain separate fail-closed
checks; successful PDF compilation is not treated as empirical validation.

\section{Label, Test, and Release Boundaries}
\label{app:boundaries}

All 300 FalsiRAG-Bench construction labels are
\texttt{machine\_seeded}. Qwen2.5 structured preannotation and deterministic
weak supervision jointly confirm 178 rows; 122 remain machine-disputed. This
is a quality audit against the same construction reference, not independent
human annotation, adjudication, or inter-annotator agreement. RAMDocs supplies
external upstream answer and document-type labels, not publication-grade human
gold for FAR's conflict and revision traces.

The 58-item FalsiRAG-Bench held-out input and RAMDocs test split remain locked
and unevaluated. Phase~B LLM jury, author adjudication, G-K/G-S, jury rescoring,
and the earlier three-system-family matrix were not run after the RAMDocs stop
rule. Completed WS2 evidence remains machine-audited development evidence;
future WS3 evidence remains public-development boundary evidence. Neither may be
described as human IAA, publication gold, or external blind testing.
