\documentclass[letterpaper]{article}
\usepackage[submission]{aaai2027}
\usepackage[hyphens]{url}
\usepackage{graphicx}
\urlstyle{rm}
\def\UrlFont{\rm}
\usepackage{natbib}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{amsmath}
\frenchspacing
\pdfinfo{/TemplateVersion (2027.1)}
\setcounter{secnumdepth}{0}

\title{Evidence Retrieved, Answers Still Wrong: Capability-Aware Failure Traces for Self-Correcting RAG}
\author{Anonymous Submission}
\affiliations{}

\begin{document}
\maketitle

\begin{abstract}
End-to-end accuracy does not reveal whether a self-correcting retrieval system failed to find evidence or failed after evidence became available. We introduce a capability-aware trace attribution protocol that compares only instrumentation shared by all methods and fails closed when a stage is not observed. Across eight methods and 350 upstream-labelled RAMDocs development items, every method has more errors after at least one correct document was retrieved and the answer text changed than retrieval-miss errors. The pooled difference is $+0.3914$ (95\% sample-cluster bootstrap $[+0.3554,+0.4275]$). This is evidence of post-retrieval answer-transformation failure, not a causal oracle decomposition: six baselines do not expose comparable detection or action traces, and two items lack upstream correct-document labels. FAR provides an instrumented case study. On a 60-item Qwen3.5 9B machine-audited development diagnostic, typed control improves over a matched untyped ablation by 0.078 (95\% paired bootstrap [0.034, 0.124]), with a positive directional reproduction across Mistral, Gemma, and Llama. A post-hoc lexical revision-delta audit exposes a whole-answer metric blind spot and reinforces mixed rather than monotone component evidence. Yet two preregistered RAMDocs rounds show no end-to-end advantage, registered clean ablations do not establish a typed-revision benefit, and a three-family ontology stress test reaches stable machine consensus on only 15/217 external items. We therefore claim a reproducible cross-method concentration of errors after retrieval and answer transformation, plus a narrower typed-control case study; we do not claim a cross-method detection bottleneck, human-validated gold, external blindness, population type mappability, or general superiority.
\end{abstract}

\section{Introduction}
Retrieval-augmented generation (RAG) grounds language models in external memory \cite{lewis2020rag}, but its default control loop is asymmetric: a query retrieves relevant passages, and generation accumulates passages that support an answer. Even iterative systems often ask whether more information is useful rather than what observation would falsify a particular claim. This asymmetry matters when the first answer has the wrong date, confuses a subsidiary with its parent, repeats an obsolete number, or upgrades correlation to causation.

We study \emph{falsification-guided retrieval}: given an initial answer, actively formulate retrievable evidence needs that could show where it is wrong. The central hypothesis is not merely that counter-evidence helps. It is that a \emph{typed evidence conflict} can serve as a control signal for both the next query and the revision action. A temporal conflict asks for version boundaries and repairs a timeline; a numerical conflict asks for alternative measurements and replaces or qualifies a value; a causal conflict seeks confounding or association-only evidence and downgrades the causal claim.

FAR makes this loop explicit. It (1) builds a claim dependency graph, (2) assigns positive typed evidence requirements, (3) generates support, refutation, and boundary queries, and (4) applies a typed revision with an auditable trace. The broader empirical problem is that methods expose different internal signals: treating an absent typed-conflict field as a failed detector would bias a cross-method analysis toward FAR. We therefore separate common observable traces from method-specific instrumentation. Our contributions are:
\begin{itemize}
    \item a preregistered, capability-aware stage-trace protocol and verifier that distinguish retrieval misses from post-retrieval unchanged and changed-answer failures without imputing unobserved detection or action states;
    \item an 8-method, 2,800-cell failure map in which post-retrieval changed-answer errors exceed retrieval misses for every method, with sample-cluster uncertainty; and
    \item an instrumented FAR case study that combines a local typed-control signal, an explicitly post-hoc lexical edit audit, directional recurrence across three additional model families, negative component evidence, two RAMDocs nulls, and a public-development applicability boundary.
\end{itemize}
The empirical claim is intentionally narrower than the architecture: the cross-method evidence locates failure after retrieval and textual answer transformation, but only the two FAR arms expose typed detection, action, and claim-level revision traces. It cannot identify a cross-method detection bottleneck or causal implementation gap.

\section{Related Work}
Active retrieval and post-hoc repair are not new. FLARE retrieves from predicted upcoming sentences when generation confidence is low \cite{jiang2023flare}; Self-RAG learns reflection tokens for retrieval and critique \cite{asai2024selfrag}; and CRAG triggers corrective retrieval after assessing retrieval quality \cite{yan2024crag}. RARR researches attribution for generated claims and edits unsupported text while preserving the draft \cite{gao2023rarr}. Chain-of-Verification asks independent verification questions \cite{dhuliawala2024cove}, while RQ-RAG learns query rewriting, decomposition, and disambiguation \cite{chan2024rqrag}.

CounterRefine is the closest direct neighbor: it explicitly treats a draft answer as a hypothesis, performs answer-conditioned counterevidence retrieval, and applies a validated KEEP/REVISE repair step \cite{huang2026counterrefine}. FAR therefore makes no priority claim for counterevidence retrieval or answer revision alone. Its narrower hypothesis is that one explicit conflict ontology should jointly select three query families and claim-specific revision actions over a dependency graph. RAMDocs and MADAM-RAG instead study ambiguity, misinformation, and noise in already retrieved conflicting documents through multi-agent debate \cite{wang2025conflicting}. FalsiRAG-Bench isolates whether typed controls retrieve and use planted counter-evidence. FEVER supplies separately licensed external annotation candidates \cite{thorne2018fever}; machine seeds are never promoted to gold.

\section{Falsification-Augmented Retrieval}
Let an initial answer $A_0$ be decomposed into claims $C=\{c_i\}_{i=1}^{n}$ and dependency edges $D$. FAR produces a revised answer $A_1$, evidence map $M$, typed conflicts $K$, and revision trace $T$.

\subsection{Claim Graph}
Each node stores text, claim type, entities, numbers, time expressions, confidence, and verifiability. An edge $(c_j,c_i)\in D$ means that $c_i$ depends on $c_j$. The implementation rejects missing dependencies and cycles and processes nodes in topological order. Structured model decomposition is accepted only if it forms a valid graph and preserves at least 80 percent of source content tokens; otherwise a conservative deterministic decomposition is used.

\subsection{Typed Evidence Requirements}
For claim $c_i$, an assigner emits requirements $R_i$. The primary requirement follows claim semantics (temporal, entity, numerical, causal, or definition), while source reliability and counter-evidence are always considered. This is a forward retrieval contract, not a label attached after retrieval.

\subsection{Three-Way Counterfactual Queries}
For each primary type $t_i$, FAR generates
\begin{equation}
Q_i=\{q_i^{+},q_i^{-},q_i^{\partial}\},
\end{equation}
where $q_i^{+}$ seeks corroboration, $q_i^{-}$ seeks an observation incompatible with the claim, and $q_i^{\partial}$ seeks scope, time, unit, definition, or comparability boundaries. Tactics depend on $t_i$: numerical refutation searches for revised values and methodologies; causal refutation searches for association-only findings, confounders, and alternative causes. Model-generated queries must contain every enabled family; invalid output falls back to deterministic typed templates.

\subsection{Typed Conflict and Revision}
Retrieved passages are compared with the claim using high-precision rules or the layered VeraRAG conflict graph. FAR maps detected relations to seven control types. A fixed priority policy chooses an action: correct time, requalify entity, replace number, downgrade causality, prefer reliable source, clarify definition, retract, or qualify uncertainty. An optional language model realizes the selected action but cannot select a different policy. Every trace records before and after text, action, conflict types, evidence IDs, rationale, and confidence. No retrieved counter-evidence is treated as absence of a detected conflict, not proof of the claim.

\section{FalsiRAG-Bench}
The machine-audited synthetic benchmark schema contains question, erroneous initial answer, atomic claims and dependencies, construction-reference evidence, counter-evidence with the refuted claim, typed conflict, expected revision, split, provenance, and annotation status.

\begin{table}[t]
\centering
\small
\begin{tabular}{lr}
\toprule
Candidate category & Count \\
\midrule
Temporal shift & 60 \\
Numerical conflict & 60 \\
Entity confusion & 60 \\
Causal overclaim & 60 \\
Multi-source conflict & 60 \\
\bottomrule
\end{tabular}
\caption{The deterministic v0.2 candidate build. Counts are construction facts, not human-validated gold statistics.}
\label{tab:data}
\end{table}

The current build derives evidence-backed candidates from VeraBench v1.1.2, with controlled perturbations and explicitly marked synthetic causal-boundary or low-reliability passages. It contains 175 documents, including 120 controlled synthetic documents. A source-document dependency group is assigned wholly to train, development, or test; the realization is 182/60/58 and has zero group leakage. The present study reports only the 60-item development split. A separate five-field test bundle is frozen for future work but is not treated as externally blind evidence.

All 300 labels are \texttt{machine\_seeded}; they are not human-validated gold. We audit them with two distinct signals that are blind to construction roles: Qwen2.5 7B structured preannotation (299/300 non-fallback outputs) and deterministic weak supervision (211/300 non-abstentions). Both signals exactly confirm the construction conflict type and action for 178 items; 122 items are disclosed as machine-disputed rather than silently relabeled. This machine agreement is quality evidence, not human inter-annotator agreement or adjudication. The construction check obtains 0.91 lexical counter-evidence recall@10 when pooling all three query families; this verifies corpus retrievability, not FAR performance.

\section{Experimental Design}
We evaluate a locally hashed Qwen3.5 9B model with matched retrieval settings. Baselines are Vanilla RAG, Multi-query RAG, Reflective RAG, closed-corpus CRAG-style and CounterRefine-style reproductions, and a Self-RAG-style inference approximation. The three style names explicitly avoid claiming equivalence to official trained or web-retrieval systems. This is a single-model development diagnostic; it does not establish multi-model generality.

Ablations remove typed conflict (untyped contradiction and generic queries), refutation queries, boundary queries, or typed revision. The primary gate compares typed and untyped FAR on aligned examples. Metrics are answer soft F1, unsupported claim rate, evidence precision, typed conflict F1, revision accuracy, overclaim reduction, and counter-evidence recall. We report category-stratified bootstrap confidence intervals, dependency-group sensitivity intervals, paired bootstrap differences, and exact McNemar tests for revision success.

\subsection{Capability-aware stage trace attribution}

We additionally reanalyse the frozen RAMDocs Round~1 artifacts for Vanilla,
Multi-Query, CRAG-style, Self-RAG-style, Reflective, CounterRefine-style, FAR,
and untyped FAR. All eight methods expose retrieved evidence identifiers, a
shared initial answer, and a final answer. Only the two FAR arms expose typed
conflicts, revision actions, and claim-level revision traces. The registered
cross-method map therefore uses only common observables and does not impute a
missing conflict field as a detector failure.

Each method--item cell is assigned once, in order, to: correct; retrieval
unscorable because no upstream correct document is labelled; retrieval miss;
post-retrieval unchanged wrong; or post-retrieval changed wrong. Answer change
is computed after removing bracketed citations and applying the frozen RAMDocs
normalizer. It denotes textual transformation, not factual correction. T1
requires changed-wrong errors to exceed retrieval misses in at least six of
eight methods. T2 bootstraps sample IDs 5,000 times, retaining all eight methods
per sampled cluster, for the pooled proportion difference. The protocol and a
second amendment for two retrieval-unscorable items were frozen before either
T1 or T2 was computed. Detection/action detail is descriptive and restricted to
the two FAR arms.

Power is treated as a gate on claim strength rather than repaired after observing results. A preregistered simulation using the Qwen discordance pattern estimates only 0.414 exact-McNemar power for the planned three-family, 60-pair-per-family reproduction at a +0.078 effect. That study is therefore designated a directional reproduction: a nonsignificant result cannot establish absence, and family-level directions must be disclosed. The same retrospective audit shows that the RAMDocs gate had a confidence-interval half-width near three percentage points, too coarse to adjudicate smaller effects.

After observing the development results, we found that whole-answer soft F1 can reward an unchanged erroneous answer because most references alter only a small date, number, entity, causal clause, or source qualifier. We therefore add a post-hoc revision-delta audit. It compares token-multiset additions and removals from the initial answer with those required by the construction reference: precision penalizes extra edits, recall penalizes retained or missing target edits, and their harmonic mean is delta F1. Typed delta F1 is zero unless the declared revision action also matches the expected action. An unchanged erroneous answer receives zero. These are lexical edit-fidelity diagnostics, not semantic correctness, and they use the frozen predictions without model calls or test access.

\begin{table}[t]
\centering
\small
\resizebox{\columnwidth}{!}{\input{../diagnostics/solo_v1/experiments/artifacts/main_results.tex}}
\caption{Machine-seeded Qwen development diagnostic ($n=60$). ``Style'' baselines are scoped reproductions. These are not blind-test or human-gold results.}
\label{tab:main}
\end{table}

\begin{table}[t]
\centering
\small
\resizebox{\columnwidth}{!}{\input{../diagnostics/solo_v1/experiments/artifacts/ablation_results.tex}}
\caption{Component ablations on the same diagnostic. The mixed results motivate our narrowed claim.}
\label{tab:ablations}
\end{table}

\section{Diagnostic Results}
FAR reaches 0.797 whole-answer correctness, above the strongest baseline value of 0.772 from the CRAG-style reproduction, and retrieves counter-evidence for 0.983 of examples (Table~\ref{tab:main}). Its absolute revision accuracy is only 0.217. The post-hoc audit gives FAR raw delta F1 0.145 (95\% stratified bootstrap [0.094, 0.208]) and typed delta F1 0.096 [0.043, 0.159]. These low absolute values confirm that the result is not evidence of a generally reliable automatic correction system.

The strongest controlled result is typed versus untyped conflict. FAR improves answer correctness by 0.078; reversing the sign reported by the ablation-centered evaluator gives a 95\% paired stratified bootstrap interval of [0.034, 0.124]. Typed-conflict F1 changes from 0 to 0.420 and revision accuracy from 0 to 0.217. Raw delta F1 rises from 0.093 to 0.145, a paired difference of +0.053 [0.003, 0.109], while typed delta F1 rises from 0 to 0.096 [0.043, 0.159]. Exact McNemar tests on revision success and typed-conflict correctness give $p=0.00024$ and $p<10^{-9}$, respectively. This supports typed conflict as a useful control signal on this diagnostic.

As a post-hoc label-audit sensitivity check, we stratify the same paired comparison by machine-consensus disposition. The answer-correctness difference remains positive on the machine-confirmed subset ($n=35$), +0.101 with 95\% bootstrap [0.039, 0.161], and on the machine-disputed subset ($n=25$), +0.047 [0.001, 0.094]. Thus the aggregate direction is not produced only by disputed labels. The subsets are small and category-imbalanced, and their dispositions are machine signals against the same construction reference; this is a sensitivity analysis, not independent label validation.

The remaining ablations contradict a stronger component-wise story. Removing refutation queries does not reduce answer correctness (0.826 versus 0.797), despite a small counter-evidence recall decrease; it also raises raw delta F1 from 0.145 to 0.194, a paired difference of +0.048 [0.012, 0.094]. Removing boundary queries is also neutral on answer correctness (0.801 versus 0.797) and on delta F1. Removing typed revision increases answer correctness to 0.873 while reducing revision accuracy and revision-action correctness to zero; its raw delta F1 falls to 0.072, a FAR-minus-ablation difference of +0.073 [0.025, 0.130]. Typed revision therefore provides explicit, more target-aligned revision behavior at a whole-answer score cost, but it is not supported as an unconditional accuracy improvement.

The unconditioned audit also prevents an overclaim in FAR's favor. CRAG-style and Vanilla obtain raw delta F1 values of 0.307 and 0.264, both above FAR's 0.145; FAR-minus-Vanilla is $-0.119$ [$-0.194,-0.038$]. Their typed delta is zero because they expose no matching typed action, but the raw result remains visible: broad answer reconstruction can cover more required lexical edits even without FAR's typed-control trace. We report all of these outcomes as negative or mixed evidence rather than attributing gains to all three query families or every repair stage.

A post-hoc frozen revision-trace audit asks whether the recorded claim-level
edits themselves align with the construction-derived target before final answer
rendering. Qwen FAR has mean trace delta F1 0.082. Only 15 of 60 traces completely
cover the construction target, while 19 are off-target and 12 make no lexical
target edit. The typed-minus-untyped trace delta is +0.048 with a paired 95\%
interval of [+0.008,+0.100], and target-complete rate rises by 0.083
[+0.017,+0.167]. However, any-target-hit rate is lower by 0.033
[$-0.117,+0.050$], so typed control does not make some target overlap more
frequent. The same post-hoc trace-F1 direction is positive in all three WS2
families, combined +0.0232 with a family-cluster interval of
[+0.0064,+0.0355]. This narrower lexical alignment signal does not establish
semantic repair: absolute fidelity remains low and collateral rewriting is
common.

\subsection{Cross-method stage trace map}

Table~\ref{tab:stage-trace} reports the registered common-observable map. T1 passes for 8/8 methods: changed-answer errors after retrieving at least one
upstream correct document exceed retrieval misses in every method. Across all
2,800 method--item cells, T2 estimates
\[
P(\text{post-retrieval changed wrong})-P(\text{retrieval miss})=+0.3914,
\]
with a 95\% sample-cluster bootstrap interval of $[+0.3554,+0.4275]$.
Retrieval miss counts range from two to six per method, whereas changed-wrong
counts range from 102 to 197. Two items have no upstream correct-document label
and contribute one retrieval-unscorable cell per method; they are not relabelled
as retrieval failures.

\begin{table*}[t]
\centering
\small
\resizebox{\textwidth}{!}{%
\begin{tabular}{lrrrrr}
\toprule
Method & Correct & Retrieval unscorable & Retrieval miss & Retrieved + unchanged wrong & Retrieved + changed wrong \\
\midrule
Vanilla RAG & 101 & 2 & 6 & 89 & 152 \\
Multi-Query RAG & 109 & 2 & 2 & 101 & 136 \\
CRAG-style & 104 & 2 & 5 & 99 & 140 \\
Self-RAG-style & 103 & 2 & 2 & 118 & 125 \\
Reflective RAG & 106 & 2 & 4 & 75 & 163 \\
CounterRefine-style & 102 & 2 & 2 & 47 & 197 \\
FAR & 109 & 2 & 3 & 134 & 102 \\
FAR minus typed conflict & 106 & 2 & 2 & 133 & 107 \\
\bottomrule
\end{tabular}
}
\caption{Capability-aware observational trace map on 350 upstream-labelled
RAMDocs development items per method. ``Changed'' means citation-stripped,
normalized answer text differs from the shared initial answer; it does not mean
the revision is factually correct.}
\label{tab:stage-trace}
\end{table*}

This map licenses a cross-method statement about where errors remain visible:
correct-document availability and answer transformation often occur without a
correct endpoint. It does \emph{not} isolate why the transformation failed.
Six baselines do not expose typed detection or action traces, so the map cannot
show that detection is sufficient, estimate a detection oracle gap, or separate
action selection from realization. Those questions require prospectively
instrumented downstream replay.

The registered post-stop-rule attribution sharpens this trade-off. Among the
226 RAMDocs items missed by both FAR and Multi-Query RAG, the earliest-failure
buckets are 4 retrieval misses, 72 undetected conflicts, 103 detected conflicts
followed by wrong revision, 35 incomplete answer sets, and 12 overfull answer
sets; the format-mismatch bucket is empty. On the 179 items with complete
correct-document retrieval, FAR remains below the baseline by 0.011 (95\% CI
$[-0.050,0.028]$). Descriptive collection F1 also changes by $-0.003$ (95\% CI
$[-0.020,0.013]$), so strict exact match is not hiding a partial-set gain.
Although conflict detection is lower on RAMDocs (0.449) than on the Qwen
diagnostic (0.883), the detected subset has only a $+0.006$ paired difference
with a confidence interval crossing zero. Accordingly, all four preregistered
mechanism hypotheses are not supported.

Within the two instrumented FAR arms, the component path analysis also rejects
our detection-only prediction. FAR has
34 positive continuous answer-score deltas over the untyped ablation, and every
one occurs on a changed-revision path; none occurs after detection without a
changed revision. Together with the higher mean obtained when typed revision is
removed, this identifies revision as a heterogeneous mediator of both local
benefit and larger aggregate harm. It does not identify a uniformly beneficial
component or rescue the RAMDocs gate.

\paragraph{Registered clean RAMDocs ablations.}
P5 reran full FAR, an aggressiveness-matched removal of typed revision, and a
flat-claim arm on all 350 upstream-labelled development items. Full exact match
is 0.3057 versus 0.3086 without typed revision; the full-minus-ablation
difference is $-0.0029$ with a 90\% sample-cluster interval of
$[-0.0229,+0.0171]$. Because that interval crosses the registered
$[-0.02,+0.02]$ equivalence boundary, H3 remains \texttt{uncertain}. The
flat-claim arm also obtains 0.3057; its difference is 0 with interval
$[-0.0057,+0.0057]$, so H5 is \texttt{equivalent} under the registered rule.
These results strengthen the negative component evidence without turning
upstream-labelled development data into human gold or a held-out result.

Table~\ref{tab:preconditions} turns these negative and mixed findings into an
operational applicability checklist. These are observed necessary checks, not
sufficient conditions or a causal theorem: satisfying one row did not by itself
restore a FAR advantage.

\begin{table}[t]
\centering
\small
\begin{tabular}{p{0.25\columnwidth}p{0.65\columnwidth}}
\toprule
Check before use & Evidence and implication \\
\midrule
Correct evidence is retrievable & Complete correct-document retrieval still gives a $-0.011$ FAR difference; reachability is necessary but insufficient. \\
Conflict is typed and detected & 72 shared errors are undetected, while the detected subset is only $+0.006$ with an interval crossing zero; detection creates an opportunity, not a guarantee. \\
Revision preserves the signal & 103 shared errors are detected then revised incorrectly. Every positive Qwen typed delta passes through a changed revision, yet removing typed revision raises the mean; revision is both mediator and risk. \\
Conclusions survive metric choice & Collection F1 changes by $-0.003$ with an interval crossing zero; partial credit must be disclosed but cannot replace the frozen gate. \\
\bottomrule
\end{tabular}
\caption{Observed applicability checks for typed conflict control. They delimit where the mechanism can be meaningfully attempted; none is sufficient for an end-to-end gain.}
\label{tab:preconditions}
\end{table}

On a visible 100-pair FEVER SUPPORTS/REFUTES transfer diagnostic, the heuristic and NLI detectors both obtain 0.72 accuracy. NLI raises conflict recall from 0.30 to 0.40, but the paired accuracy difference is 0 with a 95\% bootstrap interval of $[-0.05,0.05]$ and McNemar $p=1.0$. This does not validate typed FAR behavior outside the controlled benchmark.

\section{Mechanism Interpretation and Applicability Boundary}

\subsection{An opportunity chain, not a monotone pipeline}

Typed conflict control can influence an answer only through a sequence of
opportunities: relevant counter-evidence must be retrievable; the observed
relation must map to a usable conflict type; the detector must expose that
type; the policy must select an appropriate action; and the realized revision
must preserve correct content while changing the error. The completed evidence
does not license a multiplicative causal decomposition of these stages, but it
does show why success at one stage is not sufficient. Across all eight methods,
changed-answer errors after retrieval exceed retrieval misses; within FAR,
complete correct-document retrieval does not restore an advantage, the
detected-conflict subset is not reliably positive, and detected-then-wrong
revision is the largest shared failure bucket. The first result is a common
observable trace, while the latter results depend on FAR-specific instrumentation.
Neither is a causal oracle decomposition. Typed control is therefore best
understood as an opportunity to intervene, not as a monotone improvement applied
after retrieval.

\subsection{Why the local gain and external null can coexist}

FalsiRAG-Bench makes counter-evidence lexically reachable and constructs labels
in the same temporal, numerical, entity, causal, definition, and source-quality
vocabulary used by the controller. That alignment creates a diagnostic in
which the typed mechanism can be observed; it does not establish that the
mechanism transports. RAMDocs instead evaluates end-to-end answer-set recovery
under ambiguity, misinformation, and noise. Its two failed gates show that the
local signal is insufficient on that distribution. They do not, by themselves,
prove that construct alignment is the sole cause: the registered
H-conflict-shape test is also not supported. The defensible synthesis is thus
conditional rather than contradictory---the typed mechanism has a local signal,
while its transport conditions remain only partially identified.

\subsection{Revision is both carrier and risk}

Within the FAR case study, every observed positive Qwen typed-minus-untyped delta passes through a changed
revision, yet removing typed revision raises the aggregate answer score. This
combination rules out both a detection-only account of that local typed effect
and an always-revise policy. It does not establish that the six less-instrumented
baselines detected conflicts adequately.
It suggests a future selective-intervention design: revise only when the
expected benefit of the typed action exceeds its risk, and otherwise preserve
or abstain. No such selector has been evaluated here, so this is a design
hypothesis rather than a reported FAR capability. A proper test would register
the selector before execution and score selective risk, coverage, and answer
quality without replacing the current frozen gates.

\subsection{Cross-family directional reproduction}

WS2 tested the matched typed and untyped controls on 60 development items for
each of Mistral, Gemma, and Llama. All three family-level answer-correctness
differences are positive: $+0.0528$, $+0.0673$, and $+0.0735$, respectively.
The preregistered combined difference is $+0.0645$ over 180 pairs; the
stratified exact McNemar comparison has 31 typed-only versus 9 untyped-only
successes ($p=0.000680$), and the family-cluster bootstrap 95\% interval is
$[+0.0528,+0.0735]$. G-F therefore passes with 3/3 positive family directions.
The independently verified release remains machine-audited development
evidence. Because prospective G-P power was only 0.414, it licenses
directional reproduction across these local families, not human validation,
blind external generalization, or end-to-end superiority.

Across the three frozen WS2 families, the post-hoc raw delta difference is positive in 3/3: combined +0.0398 with a family-cluster 95\% interval of [+0.0133,+0.0536]. Typed delta is likewise 3/3 positive, combined $+0.0816$ with interval $[+0.0353,+0.1137]$. This was not a preregistered WS2 primary metric; it is a model-free sensitivity over recorded predictions, not semantic correctness or external generalization, and it does not overturn the higher raw delta of broad Qwen baselines.

\subsection{What external boundary mapping can identify}

WS3 asks whether typed effects track preregistered external conflict conditions
rather than only the construction-derived diagnostic. Its independently
verified public-development release contains 600 formal pipeline predictions
over WikiContradict and Google CONFLICTS, plus 20 calibration predictions. The
global typed-minus-untyped boundary-score differences are near zero:
$+0.0033$ on WikiContradict (95\% CI $[-0.0067,+0.0167]$) and $-0.0007$ on
Google CONFLICTS (95\% CI $[-0.0271,+0.0262]$), with Holm-adjusted McNemar
$p=1.0$ for both. These global nulls do not establish absence, because the
registered power ceiling remains low and the release is public development
evidence.

The boundary matrix is nevertheless informative. The preregistered
WikiContradict explicit and implicit predictions are contradicted: explicit
items have zero mean typed advantage, while implicit items are descriptively
larger. In contrast, the Google CONFLICTS outdated-information subgroup matches
the predicted positive direction ($+0.0040$), and the no-conflict subgroup
stays within the registered noninferiority margin ($-0.0042 \geq -0.03$). This
selects only a weak A-line reading: typed conflict control has identifiable
conditions where it can help, but the observed external effect is tiny,
dataset-dependent, and not a global transfer result. WS3 cannot rescue RAMDocs
G-A, unlock held-out data, or reinstate a general end-to-end claim.

\subsection{Machine-only ontology stability stress test}

P6-M tests whether three model families apply the frozen seven-type conflict
ontology consistently to the 217 public-development WS3 items. Each juror sees
two semantically matched views with reversed evidence order. A juror contributes
a stable vote only when its mappability label and mapped type set agree across
views; otherwise the item remains unresolved for that juror. J1, J2, and J3 are
stable on 50, 88, and 24 items, respectively. View-level mappability Fleiss
$\kappa$ is 0.167 for view A and $-0.041$ for view B.

Only 15 of 217 items reach stable machine consensus: one unanimous and 14
majority decisions; 202 are contested. This selected 6.9\% subset cannot estimate population type mappability, human agreement, or human gold. One
frozen stratum has zero consensus, so the registered six-stratum association
is not estimable. The negative result instead identifies ontology application
itself as an unstable precondition: conclusions can change under semantically
equivalent prompts and evidence order. It does not confirm H4, replace the
inactive human P6 protocol, or license extrapolation from the 15 consensus rows.

\section{Limitations and Responsible Use}
The candidate set reuses a small controlled source corpus, includes synthetic passages, and has correlated variants. Its labels are not human-validated gold, the reported split is development data, and evaluation is not externally blind. The machine audit measures agreement with construction labels and cannot reveal errors shared by construction, the LLM signal, and deterministic rules. The completed family reproduction uses four local model families on the same constructed development distribution and therefore does not establish external-domain or deployment generality. Rule conflict detectors are high precision but incomplete; model detectors inherit calibration and domain-shift risk. Retrieval failure can hide a true counterexample, while spurious low-quality contradictions can induce over-correction. The post-hoc delta metric is not semantic correctness: token multisets ignore position, valid paraphrases can be penalized, incidental copying can receive credit, and the metric inherits the construction reference. The benchmark is not current factual advice and is unsuitable for medical, legal, financial, or policy decisions. Final authors remain responsible for checking all text, citations, experiments, and disclosure requirements.

The trace audit inherits the same lexical and construction-reference limits.
It records that revision traces frequently miss the construction target or add
collateral edits, but it cannot determine whether an alternative trace is a
valid semantic paraphrase, whether the reference is uniquely correct, or which
pipeline stage causally produced the final token overlap.

The P6-M panel is a machine-sensitivity audit, not human review,
adjudication, inter-annotator agreement, or an estimate of population
mappability. The strict human P6 packet remains unanalyzed and inactive; no
model output occupies a human reviewer or adjudicator role.

The cross-method trace map is limited by heterogeneous instrumentation. It can
compare retrieved document identifiers and textual answer changes, but it cannot
infer unrecorded detector states, distinguish semantic repair from paraphrase,
or assign causal credit among detection, action, and realization. Citation-only
changes are removed before comparison, yet other meaning-preserving rewrites may
still count as changed. Two RAMDocs items have no upstream correct-document
label and are reported as retrieval-unscorable. These constraints make the map
a failure-location diagnostic rather than an oracle ladder.

Our preregistered no-budget 2+4 profile completed only its external Phase A and
then stopped. Round 1 on 350 RAMDocs development items with upstream labels gave
FAR and the strongest baseline, Multi-Query RAG, the same 0.311 strict exact
match. Round 2 changed only FAR's final answer consolidation layer; its G-A
comparison also failed, with FAR at 0.309 versus 0.311 for the frozen
Multi-Query baseline, paired difference $-0.0029$, bootstrap 95\% CI
$[-0.0314,0.0286]$, and McNemar $p=1.0$. This second failed G-A triggers the
registered stop rule and downgrades the 2+4 branch to an
applicability-boundary analysis for typed conflict control. Phase B not run:
the DeepSeek/GLM/Meta jury, author-blind adjudication, G-K/G-S, jury rescoring,
and three-family model matrix were not run. Held-out not run: neither the
FalsiRAG-Bench held-out split nor the RAMDocs held-out split was evaluated.
The RAMDocs evidence uses upstream labels and is not human inter-annotator
agreement, not publication-grade human gold, and not an externally held test.
Any future attempt to revive this profile would have to start from the recorded
failure analysis, disclose the negative refutation and boundary ablations, the
typed revision trade-off, and the FEVER null result, and demonstrate transfer
beyond the completed directional local-family reproduction before replacing
the current machine-audited development claim.

\section{Conclusion}
End-to-end accuracy hides a recurrent pattern: across eight frozen methods,
correct evidence is usually not absent when the system remains wrong, and many
errors follow an observable answer transformation. The registered trace map
finds this direction in 8/8 methods and a pooled difference of $+0.3914$, but
its capability matrix also prevents a stronger claim: six methods do not expose
comparable detection or action states. FAR turns ``what could make this answer
wrong?'' into an instrumented typed-control case study. Its local typed signal
recurs directionally across three additional families, while two RAMDocs nulls
and the registered revision ablations show that transformation is not reliably
correction. Low cross-view machine consensus further shows that stable type
mapping is itself an applicability boundary.
The resulting contribution is a reproducible account of post-retrieval answer-
transformation failure and a narrower revision-mediated FAR effect---not a
causal detection bottleneck, human-gold, blind-test, external-domain, or general
superiority claim.

\bibliography{references}
\end{document}
