\documentclass[letterpaper]{article}
\usepackage[submission]{aaai2027}
\usepackage[hyphens]{url}
\usepackage{graphicx}
\urlstyle{rm}
\def\UrlFont{\rm}
\usepackage{natbib}
\usepackage{caption}
\usepackage{booktabs}
\frenchspacing
\pdfinfo{/TemplateVersion (2027.1)}
\setcounter{secnumdepth}{0}
\title{FAR Supplementary Material}
\author{Anonymous Submission}
\affiliations{}
\begin{document}
\maketitle

\section{Schema and Machine-Audit Contract}
We provide the full machine-readable schema, deterministic build scripts,
source/corpus fingerprints, source-document-group split manifest, and masked
test inputs. A handoff command writes a gold-free directory containing only a
sanitized corpus, five-field operational test rows, and fingerprints. Test
runners do not evaluate this directory; they return unscored predictions to a
trusted scorer. The present paper profile does not claim this test as externally
blind. Machine seeds are never considered human gold. Instead, construction
labels are audited by a structured Qwen2.5 signal and deterministic weak
supervision. The artifact discloses 178 jointly confirmed and 122 disputed rows;
machine agreement is not reported as human IAA or adjudication.

\section{Implementation and Failure Handling}
Every run signature covers YAML configuration, split-specific operational
input, corpus, and Python implementation hashes. A checkpoint is flushed and
synchronized after each sample. Resume rejects signature drift. Structured LLM decomposition and query
generation are schema-validated and fall back only for malformed content;
provider failures remain errors. The typed revision policy is deterministic;
the model may realize, but not replace, its selected action.

\section{Baseline Scope}
The CRAG-style baseline grades closed-corpus retrieval and issues a corrective
query; it does not perform the original web-search extension. The Self-RAG-style
baseline approximates retrieve/relevance/support reflection at inference time
and does not use a reflection-token-trained model. Results retain these names.

\section{Statistical Protocol}
Primary intervals use 2,000 category-stratified percentile bootstrap samples
with seed 1729. System differences use paired stratified resampling over
identical IDs. Exact two-sided McNemar tests compare binary revision success.
Because candidates share source contexts, final reporting also resamples
source-document dependency groups. Undefined overclaim scores are excluded,
not treated as zeros.

\section{Executed Diagnostic Configuration}
We ran one complete inference pass per method over the same 60 development
items; there was no repeated model sampling or best-of-$n$ selection. Qwen3.5
9B ran through Ollama at temperature 0, with thinking disabled, a 1,200-token
output cap, and model unload after each sample to bound WSL memory. The host used
WSL2 with an NVIDIA RTX 4060 Laptop GPU (8,188 MiB VRAM); the generator stayed
on GPU while retrieval, reranking, and NLI ran on CPU. Retrieval used two
documents per query, BM25 ($k_1=1.5$, $b=0.75$), BGE-base-en-v1.5 dense scores,
0.4/0.6 sparse/dense fusion, and BGE-reranker-base over 20 candidates. The
NLI model was nli-distilroberta-base. Model asset revisions, the Ollama model
digest, YAML hash, implementation hash, operating-system string, predictions,
and report hashes are retained in the public diagnostic bundle. Statistical
resampling used 2,000 draws with seed 1729.

\section{Reproducibility Checklist Status}
Code, configurations, package lock, deterministic seeds, fingerprints, metric
definitions, and smoke tests are present. The reported result is explicitly a
single-model development diagnostic on construction-derived labels. No human
participant demographics, human agreement values, external blind-test result,
or multi-model claim is reported because those processes were not performed.

\end{document}
